### 2018-07-03 Thesis/Parts/Background/Background.tex
fb niels

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Introduction and Background}
\subheader{Information retrieval}
Information Retrieval(IR)is the activity of gathering relevant information, given another initial piece of information. 
\\\\The most practical example of this is a search engine. Given one or more search words the search engine will attempt to find relevant information. For example, an online search for "Information Retrieval" (initial piece of information) will give you a list of results (relevant information). To be able to do this, the search engine must know which texts are related. \\\\This can be achieved with, and without neural networks, a traditional technique which does not use a neural network is TF-IDF, shot for Term Frequency - Inverted Document Frequency. Techniques that use a neural network are newer, these are (among others) Word2Vec, Paragraph Vectors and GloVe.

\subheader{Neural Network}
A complete in-depth background into neural networks is beyond the scope of this thesis, therefore we will only describe the simplified working of a neural network and its basic application in IR.
Neural Networks are multi-layered computational software, based on matrix transformations, which tries to map input values to output values. This mapping uses a pre-defined amount of layers that modify the values through matrix transformations. Neural networks rely on training to create optimal values for these layers. Finding these optimal values is done iteratively, via a process known as "back-propogation". Back-propogation is done by starting at the output layer of the Neural Network and tracing error in the output back through the layers. This error information is used to improve the values of the layers. The values in the layers are only changed at training time; high-quality training data is therefore essential for neural networks. Neural networks can be applied to many different tasks, as long as there is sufficient training data available. The training is the process of iteratively adjusting multiplication matrices or vectors to achieve the optimal result. Which is the result that is as close as possible to the given output for all input and output sets, without under performing on other sets\footnote{Having only good results for the training/validation set is know as overfitting. This means that the layer-values are fine-tuned to only perform (extremely) well on one set, while under performing on others}.\\Neural Networks are used, in the creation of word embeddings, to either predict words that may occur around a given word or predict a word given words that surround it. For example, given the sentence\\
\begin{center}
\textit{"The quick brown fox jumps over the lazy dog"}
\end{center}
the Neural Network can be trained to either predict the words around "jumps", which we refer to as context words (the, quick, brown, fox, over, the, lazy, dog). It can also predict the word "jumps", given the context words (the, quick, brown, fox, over, the, lazy, dog). This result in a matrix, a vector per word which is referred to as a word embedding. This embedding indicates "word relatedness" which, as mentioned earlier enables association (thus retrieval) with related(relevant) texts. It is to be noted that the word embeddings are only able to represent word relatedness to the other words trained in the same run. Due to the random initialization of the initial values of the layers in the neural networks, neural networks do not produce every run the same results. Therefore, word embeddings from different runs of the same neural network cannot be compared.

@@

\subheader{Text analysis techniques}
To enable a computer to process text, for embedding creation or other tasks, the text has to be processed by an algorithm. In this research, we used embeddings created by Word2Vec and TF-IDF feature vectors.
\begin{jumpin}
\subsubheader{Word2Vec}
Word2vec word embeddings are created using a neural network, Word2vec learns word embeddings via maximizing the log conditional probability of the word given the context word(s) occurring within a fixed-sized window. Therefore the learned embeddings contain useful knowledge about word co-occurrence\cite{nalisnick2016improving}. There are multiple input/output possibilities for the neural network, best known are Skip-gram and the Continuous Bag-of-Words model (CBOW). The Skip-gram model takes a target word as input and outputs the predicted output words, while CBOW takes the context words as input and outputs the predicted target word\cite{nalisnick2016improving, pennington2014glove}. This is illustrated in Figure~\ref{figure:word2vecModels}, in this Figure, $v(w)$ represents the target word, and $v(w...)$ represent the context words. \citet{mikolov2013distributed}\cite{mikolov2013efficient} presented several extensions to the word2vec model that improve the quality of vector training and its speed, such as introducing Hierarchical Softmax, Negative Sampling and the subsampling of frequent words to the Skip-gram approach. Variations to the word2vec model have also been proposed, such as the doc2vec model described by \citet{lau2016empirical} which creates document embeddings instead of word embeddings.
\begin{center}
\begin{figure}[hbt]
\includegraphics[width=4in]{Plots/Model-Architecture-of-CBOW-and-Skip-Gram.png}
\caption{Illustration of the CBOW and Skip-Gram models, as presented by \citet{chen2017convolutional}. Left part of the Figure illustrates the CBOW model, context words are represented at the left side (input) and the center word (output) is represented as $v(w)$ as output. The right side of the Figure illustrates the Skip-Gram Model, which takes the center word as input (left) and outputs the context words (right).}\label{figure:word2vecModels}
\end{figure}
\FloatBarrier
\end{center}
\subsubheader{Paragraph vectors}
Variations on the word2vec model have also been proposed,~\citet{le2014distributed} introduced the paragraph vector in their paper "Distributed representations of sentences and documents". The Paragraph Vector framework is based on the word vectors framework. The difference between the frameworks is the calculation of the probability, the Paragraph Vector framework uses a matrix, which consists of every paragraph. This matrix is used to replace a concatenation or average of word vectors. An advantage of the paragraph vector model is that it takes the word order into consideration, at least in a small context \cite{le2014distributed}. \citet{dai2015document} state that the Paragraph Vectors model performs significantly better than other models on grouping, triplet finding and related object/article finding tasks on Wikipedia and arXiv texts.

@@

Since the word embeddings are created form a given text, these embeddings are bound to the text. All meaning embedded in the word embedding is derived from the original text. Because of this, embeddings can be "Domain-specific" meaning that it only knows words (or a specific word) in a certain context. This becomes most clear when faced with words that can have different meanings in different contexts. For this research, we categorize the embeddings into two categories, generic embeddings and domain-specific embeddings. The generic embeddings are trained on a collection of texts that use common English and contains a wide variety of topics. The domain-specific embeddings are trained on a collection of texts that uses jargon English (i.e., domain-specific terms) or is limited to a small number of topics. Given these terms, we regard the embeddings trained on the Wikipedia corpus\cite{lai2016generate, pennington2014glove, dai2015document, lau2016empirical, schnabel2015evaluation} as general, since Wikipedia uses common English and spans a wide range of topics. On the other hand, we regard the embeddings created by \citet{Truong2017Thesis} as Domain specific; these embeddings were created on academic articles, which use domain-specific terms and notations and only consists of academic texts, which contains less general/generic words compared to the Wikipedia corpus. Figure~\ref{figure:domainPlot} illustrates our corpus, marked in green, in comparison to the Wikipedia corpus, marked in blue. Our corpus is a collection of domain specific texts, written in academic language. A set that would be more specific would only focus on the individual domains or sub-domains.

@@

>>>>

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Introduction and Background}
\subheader{Information retrieval}
Information Retrieval(IR)is the activity of gathering relevant information, given another initial piece of information. 
\\\\The most practical example of this is a search engine. Given one or more search words the search engine will attempt to find relevant information. For example, an online search for "Information Retrieval" (initial piece of information) will give you a list of results (relevant information). To be able to do this, the search engine must know which texts are related. \\\\Interpreting which texts are related can be achieved with, and without neural networks, a traditional technique which does not use a neural network is TF-IDF, shot for Term Frequency - Inverted Document Frequency. Techniques that use a neural network are newer, these are (among others) Word2Vec, Paragraph Vectors and GloVe.

\subheader{Neural Network}
A complete in-depth background into neural networks is beyond the scope of this thesis, therefore we will only describe the simplified working of a neural network and its basic application in IR.
Neural Networks are multi-layered computational software, based on matrix transformations, which tries to map input values to output values. This mapping uses a pre-defined amount of layers that modify the values through matrix transformations. Neural networks rely on training to create optimal values for these layers. Finding these optimal values is done iteratively, via a process known as "back-propogation". Back-propogation is done by starting at the output layer of the Neural Network and tracing error in the output back through the layers. This error information is used to improve the values of the layers. The values in the layers are only changed at training time; high-quality (training) data is therefore essential for neural networks. Neural networks can be applied to many different tasks, as long as there is sufficient data available to create a training and validation set. The training is the process of iteratively adjusting multiplication matrices or vectors to achieve the optimal result. Which is the result that is as close as possible to the given output for all input and output sets, without under performing on other sets\footnote{Having only good results for the training/validation set is know as overfitting. This means that the layer-values are fine-tuned to only perform (extremely) well on one set, while under performing on others}.\\Neural Networks are used, in the creation of word embeddings, to either predict words that may occur around a given word or predict a word given words that surround it. For example, given the sentence\\
\begin{center}
\textit{"The quick brown fox jumps over the lazy dog"}
\end{center}
the Neural Network can be trained to either predict the words around "jumps", which we refer to as context words (the, quick, brown, fox, over, the, lazy, dog). It can also predict the word "jumps", given the context words (the, quick, brown, fox, over, the, lazy, dog). This result in a matrix, a vector per word which is referred to as a word embedding. This embedding indicates "word relatedness" which, as mentioned earlier enables association (thus retrieval) with related(relevant) texts. It is to be noted that the word embeddings are only able to represent word relatedness to the other words trained in the same run. Due to the random initialization of the initial values of the layers in the neural networks, neural networks do not produce every run the same results. Therefore, word embeddings from different runs of the same neural network cannot be compared.

@@

\subheader{Text analysis techniques}
To enable a computer to process text, for embedding creation or other tasks, the text has to be processed by an algorithm. In this research, we used embeddings created by Word2Vec and TF-IDF feature vectors.
\begin{jumpin}
\subsubheader{Word2Vec}
Word2vec word embeddings are created using a neural network, Word2Vec learns word embeddings via maximizing the log conditional probability of the word given the context word(s) occurring within a fixed-sized window. Therefore the learned embeddings contain useful knowledge about word co-occurrence\cite{nalisnick2016improving}. There are multiple input/output possibilities for the neural network, best known are Skip-gram and the Continuous Bag-of-Words model (CBOW). The Skip-gram model takes a target word as input and outputs the predicted output words, while CBOW takes the context words as input and outputs the predicted target word\cite{nalisnick2016improving, pennington2014glove}. This is illustrated in Figure~\ref{figure:word2vecModels}, in this Figure, $v(w)$ represents the target word, and $v(w...)$ represent the context words. \citet{mikolov2013distributed}\cite{mikolov2013efficient} presented several extensions to the word2vec model that improve the quality of vector training and its speed, such as introducing Hierarchical Softmax, Negative Sampling and the subsampling of frequent words to the Skip-gram approach. Variations to the word2vec model have also been proposed, such as the doc2vec model described by \citet{lau2016empirical} which creates document embeddings instead of word embeddings.
\begin{center}
\begin{figure}[hbt]
\includegraphics[width=4in]{Plots/Model-Architecture-of-CBOW-and-Skip-Gram.png}
\caption{Illustration of the CBOW and Skip-Gram models, as presented by \citet{chen2017convolutional}. Left part of the Figure illustrates the CBOW model, context words are represented at the left side (input) and the center word (output) is represented as $v(w)$ as output. The right side of the Figure illustrates the Skip-Gram Model, which takes the center word as input (left) and outputs the context words (right).}\label{figure:word2vecModels}
\end{figure}
\FloatBarrier
\end{center}
\subsubheader{Paragraph vectors}
Variations on the word2vec model have also been proposed,~\citet{le2014distributed} introduced the paragraph vector in their paper "Distributed representations of sentences and documents". The Paragraph Vector framework is based on the word vectors framework. The difference between the frameworks is the calculation of the probability, the Paragraph Vector framework uses a matrix, which consists of every paragraph. This matrix is used to replace a concatenation or average of word vectors. An advantage of the paragraph vector model is that it takes the word order into consideration, at least in a small context \cite{le2014distributed}. \citet{dai2015document} state that the Paragraph Vectors model performs significantly better than other models on grouping, triplet finding and related object/article finding tasks on Wikipedia and arXiv texts.

@@

Since the word embeddings are created from a given text, these embeddings are bound to the text. All meaning embedded in the word embedding is derived from the original text. Because of this, embeddings can be "Domain-specific" meaning that it only knows words (or a specific word) in a certain context. This becomes most clear when faced with words that can have different meanings in different contexts. For this research, we categorize the embeddings into two categories, generic embeddings and domain-specific embeddings. The generic embeddings are trained on a collection of texts that use common English and contains a wide variety of topics. The domain-specific embeddings are trained on a collection of texts that uses jargon English (i.e., domain-specific terms) or is limited to a small number of topics. Given these terms, we regard the embeddings trained on the Wikipedia corpus\cite{lai2016generate, pennington2014glove, dai2015document, lau2016empirical, schnabel2015evaluation} as general, since Wikipedia uses common English and spans a wide range of topics. On the other hand, we regard the embeddings created by \citet{Truong2017Thesis} as Domain specific; these embeddings were created on academic articles, which use domain-specific terms and notations and only consists of academic texts, which contains less general/generic words compared to the Wikipedia corpus. Figure~\ref{figure:domainPlot} illustrates our corpus, marked in green, in comparison to the Wikipedia corpus, marked in blue. Our corpus is a collection of domain specific texts, written in academic language. A set that would be more specific would only focus on the individual domains or sub-domains.

@@

\subheader{Domain-specific \& in-domain}
We defined domain specific earlier, a related term to this is "in-domain". The term domain-specific refers to a specific domain, I.E. biology or chemistry. In-domain refers to a relation between texts, this means that an article on biology is an in-domain text of another article on biology, but not on an article on chemistry. This means that two domain-specific texts do not necessarily share their domain, while an text A which is an in-domain text for text B means that text A and B share the same domain.

### 2018-07-03 Thesis/Parts/Data/Data.tex
fb niels

\citet{wiegand2018word} state that the Pareto distribution offers a good fit to the word occurrences in natural language. Their models show that, due to the additional parameters in the Pareto-III\footnote{The pareto distribution is a distribution in which recognizable by a long "tail" of low values and a quick and short ascend to the top values. More information: \url{https://en.wikipedia.org/wiki/Pareto_distribution}.}, the tail of the data fits better with the model than the Zipf\footnote{Distribution similar to the Pareto distribution, see: \url{https://en.wikipedia.org/wiki/Zipf\%27s_law}} model. This shows the relation between the word occurrences rank and the actual occurrences. This kind of word-occurrences distribution holds for many texts, including the writings of William Shakespeare and scientific texts and novels\cite{thurner2015understanding}.

@@

We split our initial set  $80\%$ - $20\%$. We use the $80\%$ set as the training set for the journal representations, and the $20\%$ set as the validation set for the journal representations. This split is based on a random number given to article each record, ensuring that all set have the same (random) training and validation set.

>>>>

\citet{wiegand2018word} state that the Pareto distribution offers a good fit to the word occurrences in natural language. Their models show that, due to the additional parameters in the Pareto-III\footnote{The pareto distribution is a distribution in which recognizable by a long "tail" of low values and a quick and short ascend to the top values. More information: \url{https://en.wikipedia.org/wiki/Pareto_distribution}.}, the tail of the data fits better with the model than the Zipf\footnote{Distribution similar to the Pareto distribution, see: \url{https://en.wikipedia.org/wiki/Zipf\%27s_law}} model. This shows the relation between the word occurrences rank and the actual occurrences. This kind of word-occurrences distribution holds for many texts, including the writings of William Shakespeare, scientific texts and novels\cite{thurner2015understanding}.

@@

We split our initial set  $80\%$ - $20\%$. We use the $80\%$ set as the training set for the journal representations, and the $20\%$ set as the validation set for the journal representations. This split is based on a random number given to each article, ensuring that all set have the same (random) training and validation set.

### 2018-07-03 Thesis/Parts/Discussion/Discussion.tex
fb niels

\subsubheader{Memory usage}
Although the TF-IDF outperforms the embeddings on the abstracts, the memory usage of the TF-IDF is higher than the memory usage of the embeddings. The top-performing embedding, TF-IDF weighted embedding, uses 3.13 GB, the top performing TF-IDF, 10K/10K uses 11.61 GB, which is 270.93\% of the storage size needed for the embedding. The closest TF-IDF configuration we used was 1K/1K, which uses 5.13 GB (as displayed in Figure~\ref{figure:tfidfPerformance}). This TF-IDF set has a median title rank of 183 and a median abstract rank of 44. Which is significantly worse than the embedding, which also uses less memory. This indicates that the embeddings are able to store the relatedness information more densely than the TF-IDF feature vectors can store the worod occurrence information.

@@

\subheader{Improvements}
This research shows that even tough the embeddings can capture and preserve relatedness, TF-IDF is able to outperform the embeddings on the abstracts. Earlier research already proposed improvements to the word embeddings. \citet{dai2015document} show that the usage of paragraph vectors improve the accuracy of word embeddings with 4.4\% on triplet creation with the Wikipedia corpus and a 3.9\% improvement on the same task based on the arXiv articles.\\
Furthermore, \citet{le2014distributed} show that the usage of paragraph vectors decrease the error rate (positive/negative) with 7.7\% compared to averaging the word embeddings on categorizing text as either positive or negative. While this looks promising, we have to keep in mind that our task differs from earlier tasks. We do not categorize on two categories but more than 3k. Still, we would expect an improvement by using paragraph vectors since the classification task if fundamentally the same, only on a much larger scale, which complicates the task due to the "grey areas"  between categories. These are the area's in which the classification algorithm is "in doubt" and could reasonably assing the article to both journals. The number of these grey areas increases given more categories. \citet{pennington2014glove} showed that the GloVe model outperforms the CBOW model, which is used in this research, on a word analogy task. \citet{wang2016linked} introduced the Linked Document Embedding method (LDE) method, which makes use of additional information about a document, such as citations. Their research specifically focused on categorizing documents, showed a 5.89\% increase of the micro-F1 score on LDE compared to CBOW, and a 9.11\% increase of the macro-F1 score. We would expect that applying this technique to our dataset would improve our scores, given earlier results on comparable tasks.\\
Even though much research has been done, we have not been able to find published results which are directly comparable to our results. This is likely due to our high amount of categorization groups, which enabled us to handle our results as a ranking problem, instead of an absolute hit, which has been used in earlier researches\cite{wang2016linked}. Even though we have an F1 score, which indicates performance on the absolute hits, our results are not comparable to other works due to the number of categories, and their overlapping subjects\footnote{As visualized in the journal embedding plots and discussed earlier}.

>>>>

\subsubheader{Memory usage}
Although the TF-IDF outperforms the embeddings on the abstracts, the memory usage of the TF-IDF is higher than the memory usage of the embeddings. The top-performing embedding, TF-IDF weighted embedding, uses 3.13 GB, the top performing TF-IDF, 10K/10K uses 11.61 GB, which is 270.93\% of the storage size needed for the embedding. The closest TF-IDF configuration (based on memory usage) we used was 1K/1K, which uses 5.13 GB (as displayed in Figure~\ref{figure:tfidfPerformance}). This TF-IDF set has a median title rank of 183 and a median abstract rank of 44. Which is significantly worse than the embedding, which also uses less memory. This indicates that the embeddings are able to store the relatedness information more densely than the TF-IDF feature vectors can store the worod occurrence information.

@@

\subheader{Improvements}
This research shows that even though the embeddings can capture and preserve relatedness, TF-IDF is able to outperform the embeddings on the abstracts. Earlier research already proposed improvements to the word embeddings. \citet{dai2015document} show that the usage of paragraph vectors improve the accuracy of word embeddings with 4.4\% on triplet creation with the Wikipedia corpus and a 3.9\% improvement on the same task based on the arXiv articles.\\
Furthermore, \citet{le2014distributed} show that the usage of paragraph vectors decrease the error rate (positive/negative) with 7.7\% compared to averaging the word embeddings on categorizing text as either positive or negative. While this looks promising, we have to keep in mind that our task differs from earlier tasks. We do not categorize on two categories but more than 3k. Still, we would expect an improvement by using paragraph vectors since the classification task if fundamentally the same, only on a much larger scale, which complicates the task due to the "grey areas"  between categories. These are the area's in which the classification algorithm is "in doubt" and could reasonably assing the article to both journals. The number of these grey areas increases given more categories. \citet{pennington2014glove} showed that the GloVe model outperforms the CBOW model, which is used in this research, on a word analogy task. \citet{wang2016linked} introduced the Linked Document Embedding method (LDE) method, which makes use of additional information about a document, such as citations. Their research specifically focused on categorizing documents, showed a 5.89\% increase of the micro-F1 score on LDE compared to CBOW, and a 9.11\% increase of the macro-F1 score. We would expect that applying this technique to our dataset would improve our scores, given earlier results on comparable tasks.\\
Even though much research has been done, we have not been able to find published results which are directly comparable to our results. This is likely due to our high amount of categorization groups, which enabled us to handle our results as a ranking problem, instead of an absolute hit, which has been used in earlier researches\cite{wang2016linked}. Even though we have an F1 score, which indicates performance on the absolute hits, our results are not comparable to other works due to the number of categories, and their overlapping subjects\footnote{As visualized in the journal embedding plots and discussed earlier}.

### 2018-07-03 Thesis/Parts/Future_work/Future_work.tex
fb niels

\subheader{Intelligent cutting}
An interesting improvement to enhance the word embeddings with could be a smarter way to remove noise, based on word embeddings. This might be achieved by analyzing the word embedding spectrum before normalization, and to then cut the center of the vector space out. This must be applied before normalization since normalization causes all embeddings to have a distance of 1 to the center point. All words which are generic are in the center of the spectrum. Removing these words prevents the larger texts to be pulled towards the middle of the vector space, where they lose the parts of their meaning which sets them apart from the other journals. We expect that this way of cutting, instead of word-occurrence cutting, will improve the quality of the word embeddings.

>>>>

\subheader{Intelligent cutting}
An interesting improvement to enhance the word embeddings with, could be a smarter way to remove noise, based on word embeddings. This might be achieved by analyzing the word embedding spectrum before normalization, and to then cut the center of the vector space out. This must be applied before normalization since normalization causes all embeddings to have a distance of 1 to the center point. All words which are generic are in the center of the spectrum. Removing these words prevents the larger texts to be pulled towards the middle of the vector space, where they lose the parts of their meaning which sets them apart from the other journals. We expect that this way of cutting, instead of word-occurrence cutting, will improve the quality of the word embeddings.

### 2018-07-03 Thesis/Parts/Motivation/Motivation.tex
fb niels

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Motivation}
\subheader{In-domain embeddings and validation}
In earlier research concerning domain specific articles,  \citet{Truong2017Thesis} found that in-domain training of the word embeddings can improve the process of document clustering. The usage of in-domain data is more important than the number of training examples and the model architecture. \citet{lai2016generate} found that the corpus domain is more important than the corpus size. Using an in-domain corpus significantly improves the performance for a given task, whereas using a corpus in an unsuitable domain may decrease performance.
Truong et al. encountered a problem in the validation of these in-domain embeddings. The word embedding produced correct document clustering results, leading to the conclusion that these embeddings are of good quality since they capture the document relatedness needed to create correct clusterings. However unpublished results by Truong et al. state that the embeddings show high error rates on the validation scores. This seems to indicate that the word-vectors are of good quality, but that the available validation metrics fail to confirm this. \citet{Truong2017Thesis} used multiple word similarity validations to asses the quality of the word embeddings. However, these sets are created to validate the generic embeddings; they fail to asses the quality of the domain-specific embeddings.

>>>>

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Motivation}
\subheader{In-domain embeddings and validation}
In earlier research concerning domain specific articles,~\citet{Truong2017Thesis} found that in-domain training of the word embeddings can improve the process of document clustering. The usage of in-domain data is more important than the number of training examples and the model architecture. \citet{lai2016generate} found that the corpus domain is more important than the corpus size. Using an in-domain corpus significantly improves the performance for a given task, whereas using a corpus in an unsuitable domain may decrease performance.
Truong et al. encountered a problem in the validation of these in-domain embeddings. The word embedding produced correct document clustering results, leading to the conclusion that these embeddings are of good quality since they capture the document relatedness needed to create correct clusterings. However unpublished results by Truong et al. state that the embeddings show high error rates on the validation scores. This seems to indicate that the word-vectors are of good quality, but that the available validation metrics fail to confirm this. \citet{Truong2017Thesis} used multiple word similarity validations to asses the quality of the word embeddings. However, these sets are created to validate the generic embeddings; they fail to asses the quality of the domain-specific embeddings.
